Using traffic analysis to identify The Second Generation Onion Router

### 2011-03-27 thesis/thesis.tex
Little bit more work in vulnerabilities

The first anonymous digital network, commonly known as MixNet was proposed in
\textcite{Chaum:1981p296}. The foundation of MixNet was the use of an
intermediary which forwarded messages from clients without any identifying
details. Numerous proposals and implementations followed such as Freedom
\parencite{Goldberg:1999p2231} and aChord \parencite{Hazel:2002p6929}. The
Second Generation Onion Router (Tor), is an anonymous network which uses a
variant of Onion Routing, to provide low latency anonymous communications to a
wide variety of Internet communications protocols
\parencite{Dingledine:2004p314}. It also aims to provide resistance against
censorship based blocking attempts \parencite{Dingledine:2008p1542}.

Tor was originally developed by the U.S. Naval Research Lab in 1997
\parencite{Syverson:1997kz} for application by the U.S. Navy. It has since been
adopted by The Tor Project, Inc. and is made publicly available, finding usage
in journalism, law enforcement, activism and military applications
\parencite{The-Tor-Project-Inc.:2011uq}. Tor is one of the most successful and
widely deployed anonymous networks, with an average of two hundred thousand
active users as of March 2011 \parencite{The-Tor-Project-Inc.:2011fk}.

Although Tor succeeds to a certain extent in providing anonymity and censorship
resistance. Under controlled conditions it is possible to distinguish Tor
traffic from other forms of encrypted traffic using traffic analysis. This makes
it easy for Tor users to be isolated and targeted for attack. Since Tor relies
on large numbers of users for the protections it provides, attacks that affect
the reliability of the network can discourage participation, making it
ineffective. This paper describes a simple traffic analysis technique which can
be used to distinguish Tor traffic from another form of commonly encrypted
traffic.

@@

While Tor has been the subject of much academic literature, the primary
objective of researchers has historically been the attempt to reveal participant
identities \parencite[3]{Murdoch:2005p325}.  Many attack techniques proposed
have been somewhat academic in nature and not necessarily feasible in practice.
Sometimes they require compromise of large parts of the Tor network, supplying
hostile data to Tor users or complicated knowledge of usage patterns and an
excess of patience. The technique demonstrated in this paper is a low cost
technique, which does not require very sophisticated equipment and can be
completed by a passive observer.

@@

Tor employs strong encryption, the goal of which is to ensure that the data
transmitted is indistinguishable from random noise. To an observer, two
different streams of encrypted data look almost the same. Thus the traditional
means of identifying these protocols are ineffective.

This raises the question: is it possible to distinguish the traffic created by
these networks, from encrypted traffic produced by conventional software? This
question can be broken down further:

\begin{enumerate*}
  \item Can the traffic be classified using automated matching techniques?
  \item Do these networks have characteristics that make them readily
    distinguishable using heuristics based matching?
  \item Do they have characteristics that make them distinguishable using
    machine learning techniques?
  %\item How long does a user have to be
  %  connected to the network before a confident match can be made?
\end{enumerate*}

\input litreview.tex

>>>>

The first anonymous digital network, commonly known as MixNet was proposed by
\citeauthor{Chaum:1981p296} in \citetitle{Chaum:1981p296}
\parencite{Chaum:1981p296}. This paper introduced a concept integral to many
future anonymity providing designs, an intermediate system responsible for
delivering messages without the identifying details of correspondents. The
intermediate system, referred to as a 'mix' also employed public key
cryptography to ensure that eavesdroppers could not obtain delivery information.

This seminal paper spurred research into new techniques for providing anonymity
and privacy for digital networks. Remailers such as Babel
\parencite{Gulcu:1996p1662} and Mixminion \parencite{Danezis:2003ys} employed a
mix to deliver anonymous email. Research produced designs that yielded
mathematically provable anonymity
\parencite{Chaum:1988p5869,Waidner:1989p5870,Berman:2004p303} and networks that
degrade gracefully when subject to attack, so called robust anonymity
\parencite{Waidner:1989p5870,Jakobsson:1998p5137}. Designs such as MorphMix
\parencite{Rennhard:2002p4559}, Tarzan \parencite{Freedman:2002kx} and P5
\parencite{P5Sherwood:2005p5872} aimed to provide anonymity features to
peer-to-peer (P2P) networks.

Providing resistance to blocking or censorship comes in many forms. The People's
Republic of China operates a sophisticated Internet firewall capable of blocking
requests to forbidden information \parencite{The-OpenNet-Initiative:2009uq}.
This system is often referred to as the Great Firewall of China. Resisting this
system can involve ignoring spurious reset packets injected into the stream of
forbidden communications \parencite{springerlink:10.1007/11957454_2}, or be as
simple as accessing data through an encrypted proxy as is done in FreeGate
\parencite{Smart:2008kx}.

More sophisticated circumvention tend to be based on anonymity providing
designs, \textcite{Kopsell:2004:ABR:1029179.1029197} describes a censorship
resistant system built on top of AN.ON
\parencite{springerlink:10.1007/3-540-44702-4_7}. Publishing systems such as
Freedom \parencite{Goldberg:1999p2231} and Freenet \parencite{Clarke:2001p2435}
aim to provide censorship resistance as well as anonymity. Infranet uses
steganography to hide sensitive messages in innocuous content Infranet
\parencite{Feamster:2002p307}.

The Second Generation Onion Router (Tor), is an anonymous network which uses a
variant of Onion Routing, to provide low latency anonymous communications to a
wide variety of Internet communications protocols
\parencite{Dingledine:2004p314}. It also aims to provide resistance against
censorship based blocking attempts \parencite{Dingledine:2008p1542}.

\section{The Second Generation Onion Router}

Onion routing was originally developed by the U.S. Naval Research Lab in 1996
\parencite{Goldschlag:1996wy} for application by the U.S. Navy. Like a mix,
messages sent over an onion routing network were encrypted with their routing
information and delivered to an intermediate server for forward delivery.
Unlike the mix however, messages delivered using the onion routing network were
encrypted multiple times, each 'layer' using a different key and routing
instructions. The first node in a chain would only be able to encrypt the
routing instructions to deliver the message to the next node. Each node in the
sequence decrypting a layer until the complete message is decrypted and
transmitted to the destination. This invention was granted a patent by the U.S.
Patent Office in 2001 \parencite{Michael:2001}.

This proposal was expanded on in \citetitle{Dingledine:2004p314} and addressed a
number of weaknesses in the original onion router proposal
\parencite{Dingledine:2004p314}. It also added some usability and anonymity
features and deviates sufficiently from the original onion routing design in
order to avoid patent claims. The technique employed by Tor is known as
telescopic onion routing.

In the Tor network, the intermediate nodes responsible for routing messages are
known as relays and are typically chained together to construct a circuit.
Traffic is repeatedly delivered through relays until it reaches another special
relay known as an exit node, where it is delivered unencrypted to the final
destination.  Figure \ref{tor-network} shows this typical usage of the Tor
network.

\begin{figure}[H]
  \centering\includegraphics[width=\linewidth]{tor-network-diagram}
  \caption{The Tor Network}
  \label{tor-network}
\end{figure}

Circuits are built incrementally, obtaining a session key from each
successive relay in the circuit and encrypting messages so that intermediate
relays are unable to decrypt the message. Since circuits can take some time to
build, they are pre-emptively created at regular intervals
\parencite[5]{Dingledine:2004p314}. Although Tor does not support many of the
traffic normalization techniques that provide strong anonymity, messages passed
through the Tor network are packaged in cells of a fixed size of 512 bytes
\parencite[5]{Dingledine:2004p314}. Tor also provides application level filters
that remove identifying features from common application protocols.

Tor is developed and maintained  by The Tor Project, Inc. and is made publicly
available, finding usage in journalism, law enforcement, activism and military
applications \parencite{The-Tor-Project-Inc.:2011uq}. Tor is one of the most
successful and widely deployed anonymous networks, with an average of two
hundred thousand active users as of March 2011
\parencite{The-Tor-Project-Inc.:2011fk}.

\section{Vulnerabilities and Weaknesses}

The Tor system was designed with a number of compromises, sacrificing some level
of anonymity in favour of usability and efficiency. This means that it is
vulnerable to attacks that aim to disrupt users, steal information or identify
users. A well known attack involves sniffing traffic that leaves exit nodes to
capture private information \parencite{website:tor-password-leak}. Many users
assume that Tor provides end to end encryption, and transmit private information
over the Tor network.

Technologies such as Javascript and Flash can be embedded in web pages accessed
by Tor users, and have control over network resources. By injecting network
traffic with certain patterns alongside regular network traffic, Tor users can
be identified \parencite{Abbott:2007p298}.

Tor bridges, intended as a way to get censored users access to Tor are easily
identified. They are also vulnerable to clogging attacks which make bridge
operators more easily identified \textcite{McLachlan:2009p197}.

Most conventional attacks against secure networks are known as traffic analysis,
this is the process of examining information about the communications rather
than the information contained within them. This information may include the
size of messages communicated, their frequency and timing. Many researchers have
proposed traffic analysis techniques that allow attackers to reveal the
identities of Tor users.

\textcite{Murdoch:2005p325} proposes a technique to identify users by estimating
the latency of individual Tor nodes using a hostile Tor node. The hostile Tor
node is able to send data to users using a predictable traffic pattern and
identify this pattern as it is repeated through the network, correlating streams
back to individuals.

This attack was shown to be impractical as the Tor network grew in size, with 
the increased number of users adding enough congestion to mask the introduced
identifying patterns \parencite{Evans:2009p315}. However a weakness in the Tor
design meant that particularly configured hostile nodes could amplify the
deliberately introduced congestion to make individuals identifiable on the
larger network.

@@

While Tor has been the subject of much academic literature, the primary
objective of researchers has historically been the attempt to reveal participant
identities \parencite[3]{Murdoch:2005p325}.  Many attack techniques proposed
have been somewhat academic in nature and not necessarily feasible in practice
\parencite{Raccoon:2008fk}. Sometimes they require compromise of large parts of
the Tor network, supplying hostile data to Tor users or complicated knowledge of
usage patterns and an excess of patience. The technique demonstrated in this
paper is a low cost technique, which does not require very sophisticated
equipment and can be completed by a passive observer.

This paper describes a simple traffic analysis technique which can be used to
distinguish Tor traffic from another form of commonly encrypted traffic.
Utilized properly this technique could be used to automatically identify Tor
nodes and block them, preventing Tor from being used as a censorship
circumvention tool.  Alternatively it could be used to isolate Tor users and
target them for attack, be it technically orientated such as a denial of service
attack or socially; coercion or arrest. Any direct attack on Tor users can
discourage participation, making the network less effective.

%Although Tor succeeds to a certain extent in providing anonymity and censorship
%resistance. Under controlled conditions it is possible to distinguish Tor
%traffic from other forms of encrypted traffic using traffic analysis. This makes
%it easy for Tor users to be isolated and targeted for attack. Since Tor relies
%on large numbers of users for the protections it provides, attacks that affect
%the reliability of the network can discourage participation, making it
%ineffective. This paper describes a simple traffic analysis technique which can
%be used to distinguish Tor traffic from another form of commonly encrypted
%traffic.

%\section{Traffic Classification}

%When considering the use of traffic analysis for classification of Internet
%communications, three techniques are used. These are: exact matching, heuristic
%matching and machine learning \parencite{Zhang:2009p1188}. Exact matching
%techniques recognise properties of communications that mirror known protocols
%and communication patterns. This might include the port an application uses or
%the format of its payload. Since Tor employs strong encryption and communicates
%using known and commonly used protocols, exact matching techniques would prove
%ineffective for classification. Heuristic matching looks for patterns in
%conversations to determine communications patterns and infer relationships,
%while machine learning involves the use of statistics to train computer
%algorithms to classify traffic. Both heuristic and machine matching techniques
%are suitable for traffic analysis of anonymous networks because they do not
%rely on analysing the content of messages.

%\subsection{Heuristic Techniques}

%The increasing burden of peer to peer (P2P) applications on campus networks,
%and their shift to encryption motivated the development of heuristics based
%techniques for identifying P2P traffic. Early techniques include identifying
%known properties of P2P networks such as often communicating using both the UDP
%and TCP protocols simultaneously and using a solitary connection to transfer
%high volumes \parencite{Karagiannis:2004p6400}. This paper also reduced the
%number of false positives by eliminating packets that matched known
%communications protocols.

%Similar techniques have been developed in \textcite{Perenyi:2006p6325} and
%\textcite{John:2008p1376}, both of which attempt to improve matching accuracy
%through early elimination of false positives and by expanding the scope of
%matching parameters. In \textcite{Oneil:2004p6451} the same heuristics are used
%to identify P2P nodes before applying a secondary heuristic to identify
%‘supernodes’ which are a subset of all P2P nodes.

%Classifying traffic based on system roles is a key feature of the technique
%that appears in \citetitle{Karagiannis:2005p6359}
%\parencite{Karagiannis:2005p6359}. By analysing communications at the host
%level then at increasing levels of granularity, a greater level of accuracy is
%achieved.

%The use of Traffic Dispersion Graphs or TDGs to identify P2P traffic appears in
%\textcite{Iliofotou:2008p6373} and \textcite{Iliofotou:2009p6461}. Once traffic
%flows are represented in a TDG, mathematical properties of the flow can be
%analysed to make positive identifications.

%Viruses have also posed a problem for network administrators, often
%over-utilizing network resources and using the network to infect new machines.
%A technique for identifying various worms was proposed in
%\textcite{Lazarevic:2003p6450}. Worms often attempt to hide their communications
%from network administrators to avoid detection, presenting the same problems of
%identification as Tor network traffic.

%\subsection{Machine Learning Techniques}

%\subsubsection{Expectation-Maximisation}

%The first use of machine learning to categorise traffic flows appears in
%\textcite{McGregor:2004p3826}. A detailed analysis of the attributes that can
%be used for machine learning and an attempt at coarse grained classification
%using an expectation-maximisation (EM) algorithm are demonstrated. The
%expectation-maximisation algorithm is a method for finding probabilistic
%estimates of parameters in a statistical model \parencite{UW:2010p7083}. The
%same technique is also demonstrated in \textcite{Soule:2004p3817} using
%histograms for finer grained classification.  The EM algorithm is again used in
%\textcite{Zander:2005p6212} and \textcite{Erman:2006p3825}, with the latter
%comparing EM favourably against a Na\"{i}ve Bayes classifier.

%\subsubsection{Na\"{i}ve Bayes}

%\textcite{Moore:2005p3827} use a supervised Na\"{i}ve Bayes to classify traffic
%flows that have previously been sorted into groups by analysing the flow
%content. This paper focuses on many of the most commonly used Internet
%protocols while \textcite{Bonfiglio:2007p6453} uses the technique for
%identifying traffic belonging to the Skype application.

%\textcite{Herrmann:2009p1189} uses Bayesian networks to fingerprint visited
%websites accessed through Privacy Enhancing Technologies (PET), including Tor.
%This technique performed poorly when applied to the Tor network, but it
%suggests that Tor traffic has particular characteristics that distinguish it
%from many existing PETs. It makes a particularly useful observation: ``The most
%frequent packet sizes in the Tor traffic dumps are, in descending order, 1500,
%−52, −638, 52, 638 and 1150 bytes, accounting for 87.6 \% of all Tor packets.''

%\subsubsection{Markov Models}

%Hidden Markov Models (HMM) are a statistical modelling technique useful
%for classification of systems with an unobservable state
%\parencite{Rabiner:1990p1153}. HMMs are first used as a traffic analysis
%technique in \citetitle{Wright:2004p3860} \parencite{Wright:2004p3860}. The
%primary identification characteristics for use with this algorithm are packet
%size and inter-packet arrival times. With refinement this algorithm is used
%with increasing accuracy in \textcite{Wright:2006p322} and
%\textcite{Dainotti:2008p1435}.

%In \textcite{Bernaille:2005p6205} a spectral clustering algorithm is used to
%discover distinguishing characteristics of traffic flows, rather than with a
%specific set of classification goals. This information is used to build HMMs
%for the purpose of traffic classification.

%Since HMMs model particular behaviours, they can be used to recognize
%deviations from normal behaviour as shown in
%\textcite{EstevezTapiador:2003p7201}. Motivated by
%\citeauthor{EstevezTapiador:2003p7201}, \textcite{Munz:2010p7085} uses TCP
%flags and the position of individual packets in a stream to train an Observable
%Markov Model for traffic classification, achieving greater classification
%accuracy than \citeauthor{Bernaille:2005p6205}.

%\subsubsection{Clustering}

%Clustering algorithms group observations into subsets based on similar
%characteristics, they are a form of unsupervised learning technique designed to
%discover organization in unlabelled data sets. They have been the basis of a
%number of traffic classification techniques with the K-Means clustering
%technique being the most prominent.  K-Means cluster analysis appears in
%\textcite{Bernaille:2006p2366}, \textcite{Erman:2007p3764} and
%\textcite{Erman:2007p6206}. The use of the Density Based Spatial Clustering of
%Applications with Noise (DBSCAN) algorithm appears in
%\textcite{Erman:2006p3766}, alongside K-Means and Autoscan algorithms.

%\subsubsection{Other Techniques}

%Other algorithms used for traffic classification include Nearest Neighbour and
%Linear Discriminant Analysis \parencite{Roughan:2004p3823}, Normalized
%Threshold \parencite{Crotti:2007p3824}. The use of the Gaussian Mixture Model
%to identify applications and identities inside SSH tunnels was demonstrated in
%\textcite{Dusi:2008p6254}. Several algorithms are compared in
%\textcite{Alshammari:2009p7474}, finding the C4.5 algorithm the most accurate
%for identifying Skype and SSH traffic.

%\subsubsection{Comparing Techniques}

%It is difficult to say what machine learning technique is the most effective as
%no consensus has been reached, the literature covers a wide variety of
%techniques each with vastly different goals and no two techniques can be
%directly compared as the data used for analysis has not been disclosed
%\parencite{Kim:2007p3867}. However some attempt has been made at comparison in
%\textcite{Williams:2006p3849} which suggests that the C4.5 algorithm has the
%greatest performance and accuracy when compared to a number of Bayesian
%algorithms. \textcite{Mohd:2009p6484} compares thirty machine learning
%algorithms  to find random tree, IBI, IBK and random forest algorithms
%obtaining the greatest classification accuracy.

%An excellent meta study on traffic analysis techniques is presented in
%\textcite{Nguyen:2008p3837}. This paper compares many of the published papers,
%the techniques used, criteria analysed and how effectively they meet stated
%goals.

@@

\begin{enumerate*}
  \item Can Tor traffic be identified using automated matching techniques?
  \item Does Tor have characteristics that make it readily distinguishable using
    heuristics based matching?
  \item Do they have characteristics that make them distinguishable using
    machine learning techniques?
\end{enumerate*}
